An Open Repository of Real-Time COVID-19 Indicators

### 2021-06-21 indicators/paper/paper.tex
Small typo fixes and corrections

% Middle authors
\author[i]{Sumit Agrawal}
\author[l]{Wael Al Saeed}
\author[g]{Taylor Arnold}
\author[j]{Amartya Basu}
\author[d]{Jacob Bien}
\author[r]{Ángel Alexander Cabrera}
\author[b]{Andrew Chin}
\author[b]{Eu Jing Chua}
\author[b]{Brian Clark}
\author[i]{Sarah Colquhoun}
\author[b]{Nat DeFries}
\author[i]{David C. Farrow}
\author[r]{Jodi Forlizzi}
\author[i]{Jed Grabman}
\author[b]{Samuel Gratzl}
\author[a]{Alden Green}
\author[r]{Robin Han}
\author[i]{Kate Harwood}
\author[a,b]{Addison J. Hu}
\author[i]{Raphael Hyde}
\author[d]{Sangwon Hyun}
\author[l]{Ananya Joshi}
\author[h]{Jimi Kim}
\author[r]{Andrew Kuznetsov}
\author[b]{Wichada La~Motte-Kerr}
\author[f]{Balasubramanian Narasimhan}
\author[r]{Jenny Yeon Jin Lee}
\author[p]{Kenneth Lee}
\author[b]{Zachary C. Lipton}
\author[r]{Michael Xieyang Liu}
\author[k]{Lester Mackey}
\author[b]{Kathryn Mazaitis}
\author[o]{Daniel J. McDonald}
\author[i]{Phillip McGuinness}
\author[i]{Michael P. O'Brien}
\author[a,b]{Natalia L. Oliveira}
\author[a,b]{Pratik Patil}
\author[r]{Adam Perer}
\author[b]{Collin A. Politsch}
\author[l]{Dawn Rucker}
\author[i]{Chris Scott}
\author[s]{Nigam Shah}
\author[n]{Vishnu Shankar}
\author[p]{James Sharpnack}
\author[b]{Dmitry Shemetov}
\author[q]{Noah Simon}
\author[i]{Benjamin Y. Smith}
\author[b]{Vishakha Srivastava}
\author[e,f]{Robert Tibshirani}
\author[e]{Elena Tuzhilina}
\author[b]{Ana Karina Van Nortwick}
\author[a]{Valérie Ventura}
\author[a]{Larry Wasserman}
\author[i]{Benjamin Weaver}
\author[m]{Jeremy C. Weiss}
\author[i]{Spencer Whitman}
\author[r]{Kristin Williams}

% Senior authors
\author[b]{Roni Rosenfeld}
\author[a,b]{Ryan J.\ Tibshirani}

@@

\paragraph{Mobility Data} SafeGraph, a company that collects geospatial data
from smartphone apps, calculates COVID-related mobility signals
\cite{SafeGraphSocial, SafeGraphPatterns} and makes it available to researchers 
under a data use agreement; we aggregate (some of) these signals to the county
level and make them publicly available.

@@

The indicators that are available in the COVIDcast API have been used in
dashboards produced by COVID Act Now \cite{CovidActNow}, COVID Exit Strategy
\cite{CovidExitStrategy}, and others; to inform the Delphi, DeepCOVID
\cite{Rodriguez:2021}, and the Institute for Health Metrics and Evaluation
(IHME) \cite{IHMEProj} COVID forecasting models; in various federal and state
government reports and analyses; and in a range of news stories. Aside from
operational use in decision-making and forecasting, they have also facilitated
numerous analyses studying the impacts of COVID-19 on the public, the
effectiveness of policy interventions, and factors that influenced the spread of
the pandemic \cite{Adjodah:2021, Pierri:2021, Jewell:2021, Chakrabarti:2020,
  Doerr:2021}. The API currently serves hundreds of thousands of requests to
thousands of users every day.

@@

\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey CLI (CTIS-CLI): The estimated percentage
  of the population with COVID-like illness based on Delphi's surveys of
  Facebook users.
\item COVID Trends and Impact Survey CLI in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know 
  someone in their local community that is sick, based on the same surveys.
\item Quidel test positivity rate (Quidel-TPR): The percentage of positive
  results among Quidel COVID antigen tests.  
\end{itemize}

@@

As a further example, Figure~\ref{fig:bexar_compare} displays data from Bexar
County, Texas (which contains San Antonio) during July 2020. On July 16, 2020,
4,810 backlogged cases were reported after reporting problems prevented them
from being reported over the past two weeks \cite{Palacios:2021}, resulting in a
clearly visible spike in the left-hand panel of Figure~\ref{fig:bexar_compare}
(case data from JHU CSSE, smoothed using a 7-day trailing average).  Meanwhile,
Delphi's COVID Trends and Impact Survey averaged around 350 responses per day in
Bexar County over the same time period, and was able to estimate the fraction of
the population who know someone in their community with COVID-Like Illness
(CLI). As we can see in the right-hand panel of Figure~\ref{fig:bexar_compare},
this signal was not affected by Bexar County's reporting problems and, as shown
in the last subsection, it is (in general) highly correlated with case rates,
providing an alternate stream of data about COVID activity unaffected by
backlogs. Similar reporting problems have occurred in many jurisdictions across
the United States, making it valuable to cross-check against external sources
not part of the same reporting systems.

@@

Third, the underlying database tracks all revisions made to the data, allowing
us to query the API to learn ``what was known when'', which is critical for
understanding the behavior (and potential pitfalls) of real-time surveillance
signals. Such revision data is rarely available in standardized format from
other sources.

>>>>

% Middle authors
\author[i]{Sumit Agrawal}
\author[l]{Wael Al Saeed}
\author[g]{Taylor Arnold}
\author[j]{Amartya Basu}
\author[d]{Jacob Bien}
\author[r]{{\'A}ngel A. Cabrera}
\author[b]{Andrew Chin}
\author[b]{Eu Jing Chua}
\author[b]{Brian Clark}
\author[i]{Sarah Colquhoun}
\author[b]{Nat DeFries}
\author[i]{David C. Farrow}
\author[r]{Jodi Forlizzi}
\author[i]{Jed Grabman}
\author[b]{Samuel Gratzl}
\author[a]{Alden Green}
\author[r]{Robin Han}
\author[i]{Kate Harwood}
\author[a,b]{Addison J. Hu}
\author[i]{Raphael Hyde}
\author[d]{Sangwon Hyun}
\author[l]{Ananya Joshi}
\author[h]{Jimi Kim}
\author[r]{Andrew Kuznetsov}
\author[b]{Wichada La Motte-Kerr}
\author[e,f]{Balasubramanian Narasimhan}
\author[r]{Yeon Jin Lee}
\author[p]{Kenneth Lee}
\author[b]{Zachary C. Lipton}
\author[r]{Michael X. Liu}
\author[k]{Lester Mackey}
\author[b]{Kathryn Mazaitis}
\author[o]{Daniel J. McDonald}
\author[i]{Phillip McGuinness}
\author[i]{Michael P. O'Brien}
\author[a,b]{Natalia L. Oliveira}
\author[a,b]{Pratik Patil}
\author[r]{Adam Perer}
\author[b]{Collin A. Politsch}
\author[l]{Dawn Rucker}
\author[i]{Chris Scott}
\author[s]{Nigam Shah}
\author[n]{Vishnu Shankar}
\author[p]{James Sharpnack}
\author[b]{Dmitry Shemetov}
\author[q]{Noah Simon}
\author[i]{Benjamin Y. Smith}
\author[b]{Vishakha Srivastava}
\author[e,f]{Robert Tibshirani}
\author[e]{Elena Tuzhilina}
\author[b]{Ana Karina Van Nortwick}
\author[a]{Val{\'e}rie Ventura}
\author[a]{Larry Wasserman}
\author[i]{Benjamin Weaver}
\author[m]{Jeremy C. Weiss}
\author[i]{Spencer Whitman}
\author[r]{Kristin Williams}

% Senior authors
\author[b]{Roni Rosenfeld}
\author[a,b]{Ryan J. Tibshirani}

@@

\paragraph{Mobility Data} SafeGraph, a company that collects geospatial data
from smartphone apps, calculates COVID-related mobility signals
\cite{SafeGraphSocial, SafeGraphPatterns} and makes them available to researchers
under a data use agreement; we aggregate (some of) these signals to the county
level and make them publicly available.

@@

The indicators that are available in the COVIDcast API have been used in
dashboards produced by COVID Act Now \cite{CovidActNow}, COVID Exit Strategy
\cite{CovidExitStrategy}, and others; to inform the Delphi, DeepCOVID
\cite{Rodriguez:2021}, and the Institute for Health Metrics and Evaluation
(IHME) \cite{IHMEProj} COVID forecasting models; in various federal and state
government reports and analyses; and in a range of news stories. Aside from
operational use in decision-making and forecasting, they have also facilitated
numerous analyses studying the impacts of COVID-19 on the public, the
effectiveness of policy interventions, and factors that influenced the spread of
the pandemic \cite{Adjodah:2021, Pierri:2021, Jewell:2021, Chakrabarti:2020,
  Doerr:2021}. The API currently serves hundreds of thousands of requests to
hundreds of users every day.

@@

\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey CLI (CTIS-CLI): The estimated percentage
  of the population with COVID-like illness based on Delphi's surveys of
  Facebook users.
\item COVID Trends and Impact Survey CLI in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know 
  someone in their local community who is sick, based on the same surveys.
\item Quidel test positivity rate (Quidel-TPR): The percentage of positive
  results among Quidel COVID antigen tests.  
\end{itemize}

@@

As a further example, Figure~\ref{fig:bexar_compare} displays data from Bexar
County, Texas (which contains San Antonio) during July 2020. On July 16, 2020,
San Antonio reported 4,810 backlogged cases after reporting problems prevented
them from being reported over the past two weeks \cite{Palacios:2021}, resulting
in a clearly visible spike in the left-hand panel of
Figure~\ref{fig:bexar_compare} (case data from JHU CSSE, smoothed using a 7-day
trailing average). Meanwhile, Delphi's COVID Trends and Impact Survey averaged
around 350 responses per day in Bexar County over the same time period, and was
able to estimate the fraction of the population who know someone in their
community with COVID-Like Illness (CLI). As we can see in the right-hand panel
of Figure~\ref{fig:bexar_compare}, this signal was not affected by Bexar
County's reporting problems and, as shown in the last subsection, it is (in
general) highly correlated with case rates, providing an alternate stream of
data about COVID activity unaffected by backlogs. Similar reporting problems
have occurred in many jurisdictions across the United States, making it valuable
to cross-check against external sources not part of the same reporting systems.

@@

Third, the underlying database tracks all revisions made to the data, allowing
us to query the API to learn ``what was known when,'' which is critical for
understanding the behavior (and potential pitfalls) of real-time surveillance
signals. Such revision data is rarely available in standardized format from
other sources.
